---
title: "HW5_statistical"
author: "Cyndi Liang"
date: "2024-02-25"
output: html_document
---

EID: 

Github Link: 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo = FALSE, include = FALSE}
library(tidyverse)
library(mosaic)
library(stringr)
library(tibble)
library(knitr)
```

# Question 1 
```{r echo = FALSE}
iron_sim = do(100000)*nflip(n=2021, prob = 0.024)
ggplot(iron_sim) + geom_histogram(aes(x=nflip), binwidth=1) + ggtitle("Distribution of Flagged Trades Simulation") + xlab("Number of Flagged Trades") + ylab("count")
iron_sum = sum(iron_sim >=70)
iron_p = iron_sum/100000
```

The histogram above displays the distribution of trades that were flagged out of 2021 total at a 2.4% probability with 100,000 monte carlo simulations. 

Null hypothesis: Iron Bank securities were flagged at the same 2.4% baseline rate than other traders. 

The test statistic was the number of trades that got flagged. The p-value is 0.00201. I don't believe the null hypothesis was plausible, as a probability of 0.201 % of having 70 or more trades flagged is way too low and seems significant in the context of this situation. 

# Question 2
```{r echo = FALSE}
bites_sim = do(100000)*nflip(n=50, prob = 0.03)
ggplot(bites_sim) + geom_histogram(aes(x=nflip), binwidth=1) + ggtitle("Distribution of Health Violations Simulation") + xlab("Number of Health Violations") + ylab("count")
bites_sum = sum(bites_sim >=8)
bites_p = bites_sum/100000
```

The histogram above displays the distribution the number of health violations issued out of 50 inspections at a 3% probability through 100,000 monte carlo simulations. 

Null hypothesis: Gourmet bites received the same frequency of health violations at a 3% chance of being flagged. 

The test statistic was the number of health violations received. The p-value is 0.00016. I don't believe the null hypothesis was plausible as at all, since the probability that 8 out of 50 inspections were considered violations was 0.016% according to the simulation, this number is most likely way too low to be considered a random variation. 

# Question 3
```{r echo = FALSE}
# part a
frequencies = read.csv("letter_frequencies.csv")
brown = readLines("brown_sentences.txt")

calculate_chi_squared = function(sentences, frequencies) {

  frequencies$Probability = frequencies$Probability / sum(frequencies$Probability)
  
  # Take out special cases and turn to uppercase
  clean_sentence = gsub("[^A-Za-z]", "", sentences)
  clean_sentence = toupper(clean_sentence)
  
  # Count letter frequencies
  observed_counts = table(factor(strsplit(clean_sentence, "")[[1]], levels = frequencies$Letter))
  
  # Calculate expected counts
  total_letters = sum(observed_counts)
  expected_counts = total_letters * frequencies$Probability
  
  chi_squared_stat = sum((observed_counts - expected_counts)^2 / expected_counts)
  
  return(chi_squared_stat)
}

calculate_all_chi_squared = function(sentences, frequencies) {
  chi_squared_results = numeric(length(sentences))
  
  for (i in seq_along(sentences)) {
    chi_squared_results[i] = calculate_chi_squared(sentences[i], frequencies)
  }
  
  return(chi_squared_results)
}

expected_chi = calculate_all_chi_squared(brown, frequencies)

```

```{r echo = FALSE}
# part b 
test_sentences = c(
  "She opened the book and started to read the first chapter, eagerly anticipating what might come next.",
  "Despite the heavy rain, they decided to go for a long walk in the park, crossing the main avenue by the fountain in the center.",
  "The museum’s new exhibit features ancient artifacts from various civilizations around the world.",
  "He carefully examined the document, looking for any clues that might help solve the mystery.",
  "The students gathered in the auditorium to listen to the guest speaker’s inspiring lecture.",
  "Feeling vexed after an arduous and zany day at work, she hoped for a peaceful and quiet evening at home, cozying up after a quick dinner with some TV, or maybe a book on her upcoming visit to Auckland.",
  "The chef demonstrated how to prepare a delicious meal using only locally sourced ingredients, focusing mainly on some excellent dinner recipes from Spain.",
  "They watched the sunset from the hilltop, marveling at the beautiful array of colors in the sky.",
  "The committee reviewed the proposal and provided many points of useful feedback to improve the project’s effectiveness.",
  "Despite the challenges faced during the project, the team worked tirelessly to ensure its successful completion, resulting in a product that exceeded everyone’s expectations."
)

test_chi = calculate_all_chi_squared(test_sentences, frequencies)

calculate_p = function(test, expected) {
  number_values_greater_equal = sum(expected >= test)
  total = length(expected)
  pvalue = number_values_greater_equal / total 
  return(pvalue)
}

p_value_results = numeric(length(test_chi))
  
for (i in seq_along(test_sentences)) {
  p_value_results[i] = calculate_p(test_chi[i], expected_chi)
}

table_part_b = tibble(
  Sentence = 1:length(p_value_results), 
  P_Value = round(p_value_results, 3)
)

pretty_table = kable(table_part_b, format = "html")
pretty_table
```


The table above displays the p_values for each of the tested sentences when compared to the expected chi square values from the brown sentences data. 

The sentence produced by a LLM should be sentence 6 because the p value for that sentence is by far the lowest out of all 10, meaning it's statistically the least likely to be written by a human. 
